% Chapter 4: Divisions & Pair Types
\chapter{Divisions \& Pair Types}
\label{ch:divisions}

Each match occurs under one of four pairing formats, defined by the \textit{type} of player and coach involved.

\section{Division Structure}

Divisions are numbered 1–3 within each class, plus a Special exhibition category:

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}llllp{4cm}@{}}
\toprule
\textbf{Division} & \textbf{Player} & \textbf{Coach} & \textbf{Primary Skill Axis} & \textbf{Example Format} \\ \midrule
IA1 / IB1 / IC1 & Human & AI & Human adaptability & "Reverse Huddle" \\[0.5em]
IA2 / IB2 / IC2 & AI & Human & Interpretability \& design & "Symbiosis Series" \\[0.5em]
IA3 / IB3 / IC3 & Human & Human & Cognitive cohesion \& clarity & "Field Vision" \\[0.5em]
Special & AI + AI & Human commentators & Style evolution & "Meta Exhibition" \\ \bottomrule
\end{tabular}
\caption{MetaSPN Division Structure}
\label{tab:divisions}
\end{table}

\section{Division 1: Human Player, AI Coach}

\subsection{Overview}
Division 1 pairs human players with AI coaching systems. This format tests human \textit{adaptability}—the ability to integrate algorithmic guidance with human intuition.

\subsection{Key Dynamics}
\begin{itemize}[leftmargin=*]
  \item AI coaches provide data-driven suggestions and pattern recognition
  \item Human players must interpret, evaluate, and apply AI recommendations
  \item Scoring emphasizes effective human-AI collaboration
\end{itemize}

\subsection{Popular Format: "Reverse Huddle"}
The AI coach presents multiple strategic options; the human player must select and justify their choice within the timeout window.

\section{Division 2: AI Player, Human Coach}

\subsection{Overview}
Division 2 inverts the relationship: AI systems execute reasoning tasks while human coaches guide strategy and meta-decisions.

\subsection{Key Dynamics}
\begin{itemize}[leftmargin=*]
  \item Human coaches shape AI behavior through prompt engineering and parameter guidance
  \item Focus on interpretability—can humans understand and guide AI reasoning?
  \item Tests human coaching skill and AI transparency
\end{itemize}

\subsection{Popular Format: "Symbiosis Series"}
Coaches compete to elicit the best performance from comparable AI systems, demonstrating coaching impact on AI capabilities.

\section{Division 3: Human Player, Human Coach}

\subsection{Overview}
Division 3 features all-human teams, emphasizing traditional coaching dynamics and interpersonal cognitive cohesion.

\subsection{Key Dynamics}
\begin{itemize}[leftmargin=*]
  \item Natural language communication and intuitive guidance
  \item Emphasis on clarity, shared mental models, and trust
  \item Showcases human collaboration at its most refined
\end{itemize}

\subsection{Popular Format: "Field Vision"}
Named for the spatial awareness in traditional sports, this format emphasizes real-time tactical communication and adaptive strategy.

\section{Special Division: AI + AI with Human Commentary}

\subsection{Overview}
The Special Division features experimental exhibitions where AI players and AI coaches compete, with human experts providing live commentary and analysis.

\subsection{Key Dynamics}
\begin{itemize}[leftmargin=*]
  \item Exploration of purely artificial cognitive systems
  \item Study of emergent coaching strategies and AI-AI interaction patterns
  \item Educational value for understanding AI capabilities and limitations
\end{itemize}

\subsection{Popular Format: "Meta Exhibition"}
High-profile matches between cutting-edge AI systems, often featuring novel architectures or training approaches, with expert human narration.

\section{Cross-Division Competition}

While divisions typically compete within their format, special \textbf{cross-division tournaments} occasionally pit different pairing types against the same challenges to compare approaches and effectiveness across collaboration modes.

