% Appendix: Glossary
\chapter{Glossary of Terms}
\label{app:glossary}

\section*{Core Concepts}

\begin{description}[style=nextline,leftmargin=0pt]

\item[Challenge]
A cognitive task or problem presented to participants during a match round.

\item[Class]
Skill tier within MetaSPN (A: Developmental, B: Performance, C: Master), determining coaching intensity and scoring weights.

\item[Coach]
A meta-agent (human or AI) that guides player strategy, decision-making, and development.

\item[Coaching Impact (CI)]
Measured influence of coaching interventions on player performance, calculated as causal delta in IdeaRank score.

\item[Division]
Match format defined by player-coach pairing type (e.g., Division 1: Human player, AI coach).

\item[Ghost Replay]
Post-match reruns conducted by human teams to verify scoring consistency and calibrate IdeaRank.

\item[IdeaRank (IR)]
The foundational algorithm for measuring thought quality across six factors: Originality, Execution, Uniqueness, Cohesion, Learning, and Quality.

\item[IdeaRank-Thought (IR-T)]
Extended scoring algorithm incorporating nine factors including correctness, compliance, and coaching impact.

\item[Meta Rating (MR)]
Composite score reflecting participant performance and consistency across matches, used for advancement decisions.

\item[MetaSPN Council]
Primary governance body responsible for rule interpretation, appeals, and strategic direction.

\item[MetaSPN Ledger]
Cryptographically secured, append-only log system recording all match data for verification and auditability.

\item[Player]
The primary participant executing reasoning tasks during a match (may be human or AI).

\item[Reason Integrity Penalty]
Score deduction applied for coaching violations or rule infractions.

\item[Reason Map Overlay]
Real-time visualization system displaying reasoning graphs, decision points, and IdeaRank factors for spectators.

\item[Referee]
Official who oversees match conduct, enforces rules, and makes real-time rulings.

\item[Round]
One of three structured segments within a match: Exploration, Execution, or Reflection.

\item[Timeout]
Designated period when coaches may communicate with players; duration and quantity vary by class.

\end{description}

\section*{Scoring Factors}

\begin{description}[style=nextline,leftmargin=0pt]

\item[Outcome Validity (O)]
Correctness and robustness of solutions or conclusions.

\item[Constraint Compliance (X)]
Adherence to rules, ethical guidelines, and resource limits.

\item[Uniqueness (U)]
Originality and creative problem framing.

\item[Cohesion (C)]
Logical and narrative consistency of reasoning.

\item[Learning (L)]
Progress, adaptation, and insight development during the match.

\item[Quality (Q)]
Clarity, communicability, and polish of presented reasoning.

\item[Trust (T)]
Verifiable grounding and testability of claims and reasoning.

\item[Density (D)]
Information efficiency and reasoning compactness.

\end{description}

\section*{Organizational Terms}

\begin{description}[style=nextline,leftmargin=0pt]

\item[Advancement Committee]
Subgroup responsible for reviewing participant eligibility for class promotion.

\item[Ethics \& Integrity Board]
Governance body overseeing fair play, AI transparency, and conduct enforcement.

\item[IdeaRank Oversight Committee (IROC)]
Technical body responsible for maintaining and calibrating the IdeaRank scoring algorithm.

\item[Master Certification]
Highest achievement level in MetaSPN (MR $\geq$ 950, 10+ Class C matches), granting governance and design privileges.

\item[Model Fingerprinting]
Process of uniquely identifying AI systems through challenge-response testing and behavioral profiling.

\end{description}

\section*{Match Types \& Formats}

\begin{description}[style=nextline,leftmargin=0pt]

\item[Audience Solve Mode]
Interactive feature allowing spectators to attempt match challenges and receive unofficial IdeaRank scores.

\item[Cross-Division Tournament]
Special competition where different pairing types (Divisions 1-3) compete on identical challenges.

\item[Exhibition Match]
Non-ranked match featuring experimental formats, celebrity participants, or showcase performances.

\item[Provisional Status]
Temporary classification for newly advanced participants (first 5 matches in new class).

\item[Reflective Timeout]
Special timeout type (Class C only) used for post-round meta-analysis rather than in-round guidance.

\item[Special Division]
Experimental format featuring AI player and AI coach with human commentary.

\end{description}

\section*{Acronyms}

\begin{description}[style=nextline,leftmargin=0pt]

\item[CI] Coaching Impact

\item[IR] IdeaRank

\item[IROC] IdeaRank Oversight Committee

\item[IRS] IdeaRank Score

\item[IR-T] IdeaRank-Thought

\item[MR] Meta Rating

\end{description}

